---
title: 'Chapter 4: Subsetting'
author: "Menglin"
date: "7/31/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 4.2 Selecting Multiple Elements

## 4.2.1 + 4.2.2 Atomic vectors and Lists

The subsetting of vectors and lists are the same. 

Positive Integer: return elements at the specificed positions
```{r}
x <- c(1,2,3,4)

x[c(3,1)]
# real number will be truncated to integer
x[c(2.6, 3.2)]
```

Negative integer: exclude elements at the specified positions
```{r}
x[-c(3,1)]
```
But we cannot mix positive and negative intergers in a single subseting operation.


Logical vector: select elements where the corresponding logical value is TRUE.
```{r}
x[c(TRUE, TRUE, FALSE, TRUE)]
```

When the original vector and the inputed index have different length, the shorter of the two is recycled to the length of the longer. But try to avoid using the recycling rules, since the rules are inconsistently applied throughtout base R which might cause unexpected result.

Nothing returen the original vector
```{r}
x[]
```

Zero returns a zero-length vector
```{r}
x[0]
```


If the vector is named, you can also use character vectors to return elements with matching names.
```{r}
y <- setNames(x, letters[1:4])
y[c("d","b","a")]
```



## 4.2.3 Matrices and Arrays

The subsetting of matrices and arrays share the same rull with subsetting of vectors and lists.

```{r}
a <- matrix(1:9, nrow = 3)
colnames(a) <- c("A", "B", "C")
a[1:2, ]
a[c(TRUE, FALSE, TRUE), c("B", "A")]
a[0,-2]
```


## 4.2.4 Data frames and tibbles

```{r}
df <- data.frame(x = 1:3, y = 3:1, z = letters[1:3])
# select the first two columns
df[1:2]

# select the first 2 rows
df[1:2, ]

# select rows based on logical vector
df[df$x == 2,]

# There are two ways to select columns from a data frame
df[c("x", "z")]
df[,c("x", "z")]
```

When you select a single column, pay special attention on how you subset it, since the outcomes are different.
```{r}
str(df["x"])

str(df[,"x"])
```

Tibble is much more consistent, since when you use [ ] for subsetting, it always returns a tubble
```{r}
df_T <- tibble::tibble(x = 1:3, y = 3:1, z = letters[1:3])

str(df_T["x"])
str(df_T[,"x"])
```

## 4.2.5 Preserving dimensionality

The default drop = TRUE behavior will simplify the returned output, i.e. it will return an object with lower dimensionality. 
```{r}
str(df[,"x", drop = FALSE])
```



# 4.3 Selecting a Single Element

There are two other subsetting operators: [[ and \$. [[ is used for extracting single items, while x$y is a useful shorthand for x[["y"]].

## 4.3.1 [[ ]]

[[]] is important when working with lists, since [] always returns a smaller list. [[]] is used to return only a single item, so you may want put a single integer or a sinlge string as the index. When a vector is inputted as index, the subsetting will be recursive, which means x[[c(1,2)]] is equivalent to x[[1]][[2]]. When you want work with single element, use [[]] can reduce the risk of getting unexpected error.

## 4.3.2 \$ 

x$y is roughly equivalent to x[["y"]]. One common mistake with \$ is to use it when you have the name of a column stored in a variable:

```{r}
var <- "x"

df$var

# Instead use [[]]
df[[var]]
```

One important difference between \$ and [[]] is \$ does partial matching.
```{r}
x <- list(abc = 1)
x$a
x[["a"]]
```

To avoid this behavior, we can set global optiuon warnPartialMatchDollar to TRUE:
```{r}
options(warnPartialMatchDollar = TRUE)
x$a
```


## 4.3.3 Missing and out-of-bounds indices

OOB: out of bound values

|row[[col]]|Zero-length|OOB(int)|OOB(chr)|Missing|
|----|----|----|----|----|
|Atomic|Error|Error|Error|Error|
|List|Error|Error|NULL|NULL|
|NULL|NULL|NULL|NULL|NULL|


When using purrr::pluck(), it will always return NULl, if inedx not exist or OOB.

When using purrr::chuck(), it will always throws an error, if index not exist or OOB.



```{r}
library(purrr)
x <- list(
  a = list(1, 2, 3),
  b = list(3, 4, 5)
)

pluck(x, "c", 1)

```


# 4.4 Subsetting and Assignment

All subsetting operators can be combined with assignment to modify selected values of an input vector: this is called subassignment. The basic form is x[i] <- value. I recommend that you should make sure that length(value) is the same as length(x[i]), and that i is unique. This is because, while R will recycle if needed, those rules are complex (particularly if i contains missing or duplicated values) and may cause problems.
Subsetting with nothing can be useful with assignment because it preserves the structure of the original object.
```{r}
df
is.data.frame(df)
df[] <- c(c(1,2,3),c(1,2,3),c(1,2,3))
is.data.frame(df)


df <- c(c(1,2,3),c(1,2,3),c(1,2,3))
is.data.frame(df)
```



























