---
title: "Advanced R Chapter 2"
author: "Menglin"
date: "7/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Advanced R
## Chapter 2: Names and Value
### 2.2 Binding basic
```{r}
library(lobstr)
library(ggplot2)
```

It’s creating an object, a vector of values, c(1, 2, 3).
And it’s binding that object to a name, x.
When refer to x, you call x and then lead to c(1,2,3)
```{r}
x <- c(1, 2, 3)
y <- x
lobstr::obj_addr(x)
lobstr::obj_addr(y)
```

lobstr::obj_addr() can show the unique identifier of c(1,2,3) in memory. As shown above, both x and y refer to the same content c(1,2,3)

### 2.3 Copy-on-modify
#### Vector
```{r}
x <- c(1, 2, 3)
y <- x

y[[3]] <- 4
x
lobstr::obj_addr(x)
lobstr::obj_addr(y)
```

x,y refer to the same contents, but when modify y, x does not change.R creat a copy of the value of x for y and modify on it. This is called copy-on-modify.

```{r}
x <- c(1, 2, 3)
cat(tracemem(x), "\n")
# tracemem() can track the memory change of some object
y <- x
y[[3]] <- 4L
y[[3]] <- 5L
x[[3]] <- 4L
untracemem(x)
```
Every modification on y will make a new copy of value and apply the modification on it. the origional value, which was binding on x, will not be copied when being modified.

#### Function
```{r}
f <- function(a) {
  a
}

x <- c(1, 2, 3)
cat(tracemem(x), "\n")
#> <0x3df1848>

z <- f(x)
# there's no copy here!

untracemem(x)
```

While f() is running, the a inside the function points to the same value as the x does outside the function. After runing f(), z and x will point to the same value. Since this value is not modified, no copy was made.

#### List
```{r}
l1 <- list(1, 2, 3)
l2 <- l1
l2[[3]] <- 4
ref(l1, l2)
a <- list(x,y)
ref(a)
```

lobstr::ref() prints the memory address of each object, along with a local ID
list is different from vector. Vector is value in memory, but list is a relation of several elementary value.

#### Data Frame
```{r}
d1 <- data.frame(x = c(1, 5, 6), y = c(2, 4, 3))
d2 <- d1
d2[, 2] <- d2[, 2] * 2
d3 <- d1
d3[1, ] <- d3[1, ] * 3
ref(d1,d2,d3)
```


Data frames are lists of vectors, and it also use copy-on-modify behavior.

#### Character Vector
```{r}
x <- c("a", "a", "abc", "d")
ref(x, character = TRUE)
```

### 2.4 Object Size
You can find out how much memory an object takes with lobstr::obj_size().
```{r}

obj_size(letters)
obj_size(ggplot2::diamonds)
```
```{r}
x <- runif(1e6)
obj_size(x)
```
```{r}
y <- list(x, x, x)
obj_size(y)
```

Finally, R 3.5.0 and later versions have a feature that might lead to surprises: ALTREP, short for alternative representation. This allows R to represent certain types of vectors very compactly. The place you are most likely to see this is with : because instead of storing every single number in the sequence, R just stores the first and last number. This means that every sequence, no matter how large, is the same size:

```{r}
obj_size(1:3)
obj_size(1:1e3)
obj_size(1:1e6)
```

### 2.5 Modify in place (Not Really)
```{r}
v <- c(1, 2, 3)
ref(v)
tracemem(v)
v[[3]] <- 4
ref(v)
untracemem(v)
```
When value has only one name, when modify the value, R still make a copy.
如果运行对性能运行有要求，直接用tracemen()来观察内存情况。

#### Environment
```{r}
e1 <- rlang::env(a = 1, b = 2, c = 3)
e2 <- e1
ref(e2,e1)

e2$c <- 4
ref(e2)
```
environment always modify in place

```{r}
e <- rlang::env()
e$self <- e

ref(e)
```

Environment can contain itself.

### 2.6 Unbinding and garbage collector
Both R and the operating system are lazy: they won’t reclaim memory until it’s actually needed.Looking from the outside, it’s basically impossible to predict when the GC will run. In fact, you shouldn’t even try. If you want to find out when the GC runs, call gcinfo(TRUE) and GC will print a message to the console every time it runs.

You can force garbage collection by calling gc(). But despite what you might have read elsewhere, there’s never any need to call gc() yourself. The only reasons you might want to call gc() is to ask R to return memory to your operating system so other programs can use it, or for the side-effect that tells you how much memory is currently being used.lobstr::mem_used() is a wrapper around gc() that prints the total number of bytes used.
```{r}
mem_used()
```


































